---
title: "Assignment_3"
author: "HEEREKAR MADHUKAR"
date: '2022-10-18'
output:
  pdf_document: default
  word_document: default
---
 
1. Formulate and solve this transportation problem using R

Converting all details into table format

```{r}
tab <- matrix(c(22,14,30,600,100,
                16,20,24,625,120,
                80,60,70,"-","-") , ncol=5 , byrow=TRUE)

colnames(tab) <- c("Warehouse1","Warehouse2","Warehouse3","Prod cost","Prod Capacity") 
row.names(tab) <- c("Plant A","Plant B","Demand")
tab <- as.table(tab)
tab
```
$$\text{Min} \hspace{.2cm} TC = 622 X_{11} +614 X_{12} +630 X_{13} +641 X_{21}+ 645X_{22}+649X_{23}$$ 
 

/text{subject to}

#Production Capacity consntraints 
Production plant A :
$$ X_{11} +X_{12}+X_{13}+ \le 100 $$
Production Plant B :
$$ X_{21} +X_{22}+X_{23}+ \le 120 $$
#Demand Consntraints 

Demand Warehouse 1 :
$$ X_{11} + X_{21} \ge 80 $$
Demand Warehouse 2 :
$$ X_{12} + X_{22} \ge 60 $$
Demand Warehouse 3 :
$$ X_{13} + X_{23} \ge 70 $$
Non-negativity of the variables

$$ X_{ij} \ge 0 $$
Where 

$$ i = 1,2,3 $$
And 

$$ j = 1,2,3 $$
Since demand and supply are not equal, the system is out of balance, so we constructed the dummy row warehouse 4.

```{r}
library(lpSolveAPI)
library(lpSolve)
# Set up cost matrix
Transportcost <- matrix(c(622,614,630,0,
                641,645,649,0) , ncol=4 , byrow=TRUE)
#defyning rows and coloumns 
colnames(Transportcost) <- c("Warehouse_1","Warehouse_2","Warehouse_3","Dummy")
rownames(Transportcost) <- c("Plant_A", "Plant_B")
Transportcost
```
```{r}
#setting up constraint signs and right-hand sides(Production side)
row.signs <- rep("<=",2)
row.rhs <- c(100,120)

#Demand side constraints#
col.signs <- rep(">=",4)
col.rhs <- c(80,60,70,10)

#solve the model
lptrans <- lp.transport(Transportcost,"min",row.signs,row.rhs,col.signs,col.rhs)

lptrans$solution
```
I obtained the variables values after solving the transportation issue as
$$x_{12}= 60$$
$$x_{13}= 40$$
$$x_{21}= 80$$
$$x_{23}= 30$$
$$x_{24}= 10$$
```{r}
lptrans$objval
```
2)Formulate the dual of the transportation problem

As we all know, the first priority was to reduce transportation costs, and the second priority would be to increase value added (VA).

```{r}
cost_2 <- matrix(c(622,614,630,100,"u1",
641,645,649,120,"u2",
80,60,70,220,"-",
"v1","v2","v3","-","-"),ncol = 5,nrow = 4,byrow = TRUE)
colnames(cost_2) <- c("Warehouse_1", "Warehouse_2","Warehouse_3","Production Capacity","Supply(Dual)")
rownames(cost_2) <- c("Plant_A","Plant_B","Demand","Demand(Dual)")
```


p and  q will be the variables for the dual.
$$  \text{Max} \hspace{.2cm} Z = 100 p_{1} + 120 p_{2} +80 q_{1} + 60 q_{2} + 70 q_{3}$$


Subject to the following constraints

$$  p_{1} + q_{1} \le 622 $$ 
$$  p_{1} + q_{2} \le 614 $$ 
$$  p_{1} + q_{3} \le 630 $$ 
$$  p_{2} + q_{1} \le 641 $$ 

$$  p_{2} + q_{2} \le 645 $$ 
$$  p_{2} + q_{3} \le 649 $$ 
Where
y1 = Warehouse_1

y2 = Warehouse_2

y3 = Warehouse_3

x1 = Plant_1

x2 = Plant_2
```{r}
#Objective function

f.obj <- c(100,120,80,60,70)

#transposed from the constraints matrix in the primal
f.con <- matrix(c(1,0,1,0,0,
                  1,0,0,1,0,
                  1,0,0,0,1,
                  0,1,1,0,0,
                  0,1,0,1,0,
                  0,1,0,0,1), nrow = 6, byrow = TRUE)

f.dir <- c("<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=")

f.rhs <- c(622,614,630,641,645,649)
lp("max",f.obj,f.con,f.dir,f.rhs)

lp("max",f.obj,f.con,f.dir,f.rhs)$solution
```
So Z=139,120 dollars and variables are: $$ p_{1} = 614 $$which represents Plant A
 $$ p_{2} = 633 $$which represents Plant B
$$ q_{1} = 8 $$ which represents Warehouse 1
 $$ q_{3} = 16 $$which represents Warehouse 3
 
3) Economic Interpretation of the dual

Observations:

Using the available data and restrictions, the maximum shipping and production expenses will be 139,120 dollars.

Z ranges from 132790 (Primal) to 139120 (Maximum) (Dual). The goal of this issue is to identify a maximum and a minimum. As a result, we realized that we shouldn't be shipping simultaneously from Plant(A/B) to all three warehouses. From where we should be shipping:

$$60 p_{12}$$ which is 60 Units from Plant A to Warehouse 2.


$$40 p_{13}$$ which is 40 Units from Plant A to Warehouse 3.

$$80 p_{21}$$which is 80 Units from Plant B to Warehouse 1.

$$30 p_{23}$$ which is 30 Units from Plant B to Warehouse 3.
We will Max the profit from each distribution to the respective capacity.

We have the following:
$$  p_{1} ^0 -y_{1} ^0 \le 622 $$ then we subtract $$  q_{1} ^0 $$ to the other side to get  $$ p_{1} ^0\le 622 - q_{1} ^0 $$
To compute it would be $614 <= (-8+622) which is correct. we would continue to evaluate these equations:
$$  p_{1}\le 622 - q_{1} => 614\le622 - 8 = 614=>correct$$
$$  p_{1}\le 614 - q_{2} => 614\le614 - 0 = 614=>correct$$
$$  p_{1}\le 630 - q_{3} => 614\le630 - 16 = 614=>correct$$
$$  p_{2}\le 641 - q_{1} => 633\le614 - 8 = 633=>correct$$
$$  p_{2}\le 645 - q_{2} => 633\le645 - 0 = 645=> Incorrect$$
$$  p_{2}\le 649 - q_{3} => 633\le649 - 16 = 633=>correct$$
By updating each of the columns, we may test for the shadow price after learning from the Duality-and-Sensitivity. We swap out 100 and 120 in our LP Transportation issue for 101 and 121, respectively. R is seen here.

```{r}
row.rhs1 <- c(101,120)
row.signs1 <- rep("<=",2)
col.rhs1 <- c(80,60,70,10)
col.signs1 <- rep(">=",4)
row.rhs2 <- c(100,121)
row.signs2 <- rep("<=",2)
col.rhs2 <- c(80,60,70,10)
col.signs2 <- rep(">=",4)

lp.transport(Transportcost,"min",row.signs,row.rhs,col.signs,col.rhs)

lp.transport(Transportcost,"min",row.signs1,row.rhs1,col.signs1,col.rhs1)
lp.transport(Transportcost,"min",row.signs2,row.rhs2,col.signs2,col.rhs2)
```

By choosing the minimum of this particular function, the number decreasing by 19 indicates that the shadow price, which was determined by adding 1 to each plant, is 19. There isn't a shadow price for the Plant B.

From the dual variable $$ q_{2}$$ where Marginal Revenue <= Marginal Cost. The equation was $$  p_{2}\le 645 - q_{2} => 633\le645 - 0 = 645=> Incorrect$$ and this was found by using $$  p_{1} ^0 - q_{1} ^0 \le 622 $$ then we subtract $$  q_{1} ^0 $$ to the other side to get  $$ p_{1} ^0\le 622 - q_{1} ^0 $$

```{r}
lp("max", f.obj,f.con, f.dir,f.rhs)$solution
```

$$q_{2}=0$$ .

The interpretation from above:
from the primal:
 $$60p_{12}$$which is 60 Units from Plant A to Warehouse 2.
 $$40p_{13}$$which is 40 Units from Plant A to Warehouse 3.
 $$80p_{21}$$which is 80 Units from Plant B to Warehouse 1.
 $$30p_{23}$$which is 60 Units from Plant B to Warehouse 3.
 
from the dual

Our aim is to get MR=MC. MR = MC in five of the six instances. The only plant that does not meet this condition is Plant B to Warehouse 2. We can see from the primal that no AEDs will be sent there.